---
title: "House Prices in New York"
author: "Group CC901E2: SID's"
institute: "The University of Sydney"
format:
  revealjs:
    self-contained: true
    fig-format: retina
    toc: true
    toc-depth: 1
    number-sections: false
    code-line-numbers: false
    slide-number: c
    scrollable: true
    pdf-max-pages-per-slide: 1
    history: false # don't add each slide to browser history
    
execute:
  echo: false
  warning: false # do not show warning in the chunk
  message: false # do not show message in the chunk
---

# Introduction

```{r message=FALSE}
# Load packages here
library(tidyverse)
library(ggplot2)
library(ggfortify)
library(dplyr)
library(ISLR)
library(sjPlot)
library(leaps)
library(GGally)
library(caret)
library(mplot)
```

```{r}
# The original data is loaded here
#| output: FALSE
house = read.delim("housing-prices-ge19.txt", 
                    header = TRUE)
# convert data types
house = house %>% mutate (
  Waterfront = as.factor(Waterfront),
  New.Construct = as.factor(New.Construct),
  Central.Air = as.factor(Central.Air),
  Fuel.Type = as.factor(Fuel.Type),
  Heat.Type = as.factor(Heat.Type),
  Sewer.Type = as.factor(Sewer.Type),
  Test = as.factor(Test)
)

# Remove the Test column, because we do not have the metadata
house = house %>%
  select(-Test)

# clean names
# house = janitor::clean_names(house)
```




# More Discussion

Model stability

## `mplot` package {.smaller}

It takes some time to create the `vis` and `af` objects for plotting variable inclusion plot and model stability plot, so we can save it as `RData`.

```{r, echo=TRUE, eval=FALSE}
vis.h = mplot::vis(lm.h.f, B = 100, seed = 2022)
af.h <- mplot::af(lm.h.f, B = 100, n.c = 300, seed = 2022)
save(vis.h, af.h, file = "house_main.RData")
```
Then, load it when we need.
```{r, echo=TRUE}
load("house_main.RData")
```

## Variable Inclusion Plot {.smaller}

```{r echo=TRUE, fig.fullwidth=TRUE}
plot(vis.h, which = "vip", interactive = FALSE, legend.position = "bottom")
```

## Variable Inclusion Plot: Insight {.smaller}

-   `Land.Value`, `Living.Area`, `Bathrooms`, `New.Construct`, `Waterfront` are the five most important variables for predicting `Price`.
-   The non-monotonic nature of the `HeatTypeHot.Air` line indicates that a group of variables contains similar information to it.
-   The path of `Sewer.TypePublic` and all levels of `FuelType` lie below the path of redundant variable, which means they are included in models by chance (don't provide any useful information).

## Model stability Plot {.smaller}

-   There appears to be dominant models in models of size **two, three, and five** (including the intercept), as demonstrated by one of the circles being substantially larger than the other circles with models of the same size.
-   The stepwise model (size = 14) is not stable as all circles of that size are small.

```{r, echo=TRUE}
plot(vis.h, highlight = "Heat.TypeHot.Air", which = "boot", interaction = F, max.circle = 10, seed = 2022, legend.position = "bottom")
```

## Which models are dominant? {.smaller}

```{r, echo=TRUE}
print(vis.h, min.prob = 0.5)
```
- The simple regression model $\{Living.Area\}$ and $\{Land.Value, Living.Area\}$ model are selected in all bootstrap resamples.
- The $\{Waterfront1, Land.Value, Living.Area, Bathrooms\}$ model is selected in 78% of bootstrap resamples.
