---
title: "R Notebook"
output: html_notebook
---

page3, simple regression
* independence analysis are needed in the previous page, as it discussed for meta data*

```{r "setup"}
#| message: FALSE
library(tidyverse)
library(ggfortify)
library(GGally)

```

```{r "read-data"}
#| output: FALSE
house = read.delim("housing-prices-ge19.txt", 
                    header = TRUE)
# convert data types
house = house %>% mutate (
  Waterfront = as.factor(Waterfront),
  New.Construct = as.factor(New.Construct),
  Central.Air = as.factor(Central.Air),
  Fuel.Type = as.factor(Fuel.Type),
  Heat.Type = as.factor(Heat.Type),
  Sewer.Type = as.factor(Sewer.Type)
)

house
# the previous is using Muwei made in contract, can be just use yours to combine.
```

```{r}
# difference begins
house_num = subset(house,select = c(Price, Lot.Size, Age, Land.Value,Living.Area, Pct.College, Bedrooms, Fireplaces, Bathrooms, Rooms))
house_num
```

```{r}
# GGally::ggpairs(house_num)
# a little bit messy
```

```{r}
cor_house = cor(house_num)

melted_cor_house = cor_house %>%
 data.frame() %>%
 rownames_to_column(var = "var1") %>%
 gather(key = "var2", value = "cor", -var1)
```

```{r}
ggplot(data = melted_cor_house, 
 aes(x=var1, y=var2, fill=cor)) +
 geom_tile() + theme_minimal(base_size = 22)+
 scale_fill_gradient2(
 low = "blue", high = "red", mid = "white",
 midpoint = 0, limit = c(-1,1)) +
 theme(axis.text.x = element_text(angle = 90))
```

```{r}
qtlcharts::iplotCorr(house_num)
```

Since our depend variable is 'Price', based on these two graph, it can easily found that the variable 'Living.Area' may be the most affected by 'Price' (0.71, with dark pink).

### Fitting a straight line by least squares

dependent variable: Price

independent variable: Living.Area

```{r}
lm1 = lm(Price ~ Living.Area, 
 data = house)
lm1
```

### Checking for Assumption:

**Assumptions:** The residuals $\varepsilon_i$are iid${\mathcal N}(0,\sigma^2)$ and there is a linear relationship between y and x.

```{r}
library(ggfortify)
autoplot(lm1, which = 1:2) + theme_bw()
```

-   Linearity: The Auxiliary line is reasonably well plotted like a straight line (with no obvious curve), so there is no obvious pattern in the residual vs fitted values plot.

-   Homoskedasticity: It appears the residuals are getting spread-out and do not appear to be fanning out or changing their variability over the range of the fitted values so the constant error variance assumption is met.

-   Normality: in the QQ plot, the points are reasonably close to the diagonal line. Although there seems to be some outliers exist, the data-set is relatively large(with 1733 observations). Thus, the normality assumption is approximately satisfied.

```{r}
plot(Price ~ Living.Area, 
 data = house)
abline(lm1, lwd = 3, col = "red")
```

### Conclusion：
As the assumption all met, it can be concluded that our simple estimated model is $\hat{Price}$ = 12796.6 + 113.4 Living area
