---
title: House Prices in New York

# Use letters for affiliations
author:
  - name: CC901E2
    affiliation: a
    
address:
  - code: a
    address: The University of Sydney

# Optional: line of arbitrary text with additional information.
# Could be used, for example, to mention the bibliographic info in a post-print.
# If not specified, defaults to "This version was compiled on \today"
date_subtitle: \today

# For footer text  TODO(fold into template, allow free form two-authors)
lead_author_surname: CC901E2

# Place eg a DOI URL or CRAN Package URL here
#doi_footer: "https://cran.r-project.org/package=YourPackage"

# Abstract
abstract: |
  Abstract here
  
# # Optional: Acknowledgements
# acknowledgements: |
#   This template package builds upon, and extends, the work of the excellent
#   [rticles](https://cran.r-project.org/package=rticles) package, and both packages rely on the
#   [PNAS LaTeX](http://www.pnas.org/site/authors/latex.xhtml) macros. Both these sources are
#   gratefully acknowledged as this work would not have been possible without them.  Our extensions
#   are under the same respective licensing term
#   ([GPL-3](https://www.gnu.org/licenses/gpl-3.0.en.html) and
#   [LPPL (>= 1.3)](https://www.latex-project.org/lppl/)).
# # Optional: One or more keywords

keywords:
  - keywords1
  - keywords2
  - keywords3
  - keywords4
  - keywords5

# Paper size for the document, values of letter and a4
papersize: letter

# Font size of the document, values of 9pt (default), 10pt, 11pt and 12pt
fontsize: 9pt

# Optional: Force one-column layout, default is two-column
#one_column: true

# Optional: Enables lineno mode, but only if one_column mode is also true
#lineno: true

# Optional: Enable one-sided layout, default is two-sided
#one_sided: true

# Optional: Enable section numbering, default is unnumbered
numbersections: true

# Optional: Specify the depth of section number, default is 5
#secnumdepth: 5

# Optional: Skip inserting final break between acknowledgements, default is false
skip_final_break: true

# Optional: Bibliography 
bibliography: ref

# Optional: Enable a 'Draft' watermark on the document
#watermark: true

# Customize footer, eg by referencing the vignette
footer_contents: "DATA 2902: Executive Summary Report"

# Produce a pinp document
# output: 
#     pinp::pinp

output:
  pinp::pinp
    

# # Required: Vignette metadata for inclusion in a package.
# vignette: >
#   %\VignetteIndexEntry{YourPackage-vignetteentry}
#   %\VignetteKeywords{YourPackage, r, anotherkeyword}
#   %\VignettePackage{YourPackage}
#   %\VignetteEngine{knitr::rmarkdown}
---

```{r message=FALSE, warning = FALSE, echo = FALSE}
# Hide all the codes
knitr::opts_chunk$set(echo=FALSE)
```

```{r message=FALSE, warning = FALSE}
# Load packages here
library(tidyverse)
library(ggplot2)
library(ggfortify)
library(dplyr)
library(ISLR)
library(sjPlot)
library(leaps)
library(GGally)
library(caret)
library(mplot)
library(lmSubsets)
```

```{r}
# The original data is loaded here
#| output: FALSE
house = read.delim("housing-prices-ge19.txt", 
                    header = TRUE)
# convert data types
house = house %>% mutate (
  Waterfront = as.factor(Waterfront),
  New.Construct = as.factor(New.Construct),
  Central.Air = as.factor(Central.Air),
  Fuel.Type = as.factor(Fuel.Type),
  Heat.Type = as.factor(Heat.Type),
  Sewer.Type = as.factor(Sewer.Type),
  Test = as.factor(Test)
)

# Remove the Test column, because we do not have the metadata
house = house %>%
  select(-Test)
```


# Introduction 
Introduction here

To cite, edit the ref.bib document; and use \cite{PeerJ:Rcpp} here.

Once you cited, the references will automatically show up at the end of the article.


## two "#" will make a 'subtitle'
Introduction subsection if needed



# Data set description
Data description here




# Analysis {#sec-analysis}


# Model Stability {#sec-stability}


# Model Evaluaion and Results

```{r}
# backward AIC

# Sample Mean of Price
M0_back = lm(house$Price ~ 1, data = house)
# Price against everything except itself
M1_back = lm(house$Price ~ ., data = house)

# Dropping of variables backwards until lowest AIC is reached
step.back.aic = step(M1_back,
                     direction = "backward",
                     trace = FALSE)

```


```{r}
# Forward AIC

# Sample Mean of Price
M0_fwd = lm(house$Price ~ 1, data = house)
# Price against everything except itself
M1_fwd = lm(house$Price ~ ., data = house)

# Dropping of variables forwards until lowest AIC is reached
step.fwd.aic = step(M0_fwd, scope = list(lower = M0_fwd, upper = M1_fwd),
                     direction = "forward",
                     trace = FALSE)
```


```{r}
# Stable model
stable_model = lm(Price ~ Waterfront + Land.Value + Living.Area + Bathrooms, data = house)
```

```{r}
# in-sample performance

# r^2 and adjusted r^2 of backward
rsquare_back = summary(step.back.aic)$r.square
rsquare_adj_back = summary(step.back.aic)$adj.r.square

# r^2 and adjusted r^2 of forward
rsquare_fwd = summary(step.fwd.aic)$r.square
rsquare_adj_fwd = summary(step.fwd.aic)$adj.r.square

# r^2 and adjusted r^2 of stable
rsquare_stable = summary(stable_model)$r.square
rsquare_adj_stable = summary(stable_model)$adj.r.square

# make a table
models = c("Backward", "Forward", "Stable")
r_square = c(rsquare_back, rsquare_fwd, rsquare_stable)
r_adj_square = c(rsquare_adj_back, rsquare_adj_fwd, rsquare_adj_stable)

rsquare_table <- data.frame(models, r_square, r_adj_square)

## round into three decimal
rsquare_table = rsquare_table %>%
  mutate_if(is.numeric, 
            round, 
            digits = 3)

```

```{r}
set.seed(2022)

# set 10 folds cross validation
tr_ctrl = trainControl(
    method = "cv", number = 10,
    verboseIter = FALSE
)

# backward
cv_back = train(
        Price ~ Lot.Size + Waterfront + Age + Land.Value + 
                New.Construct + Central.Air + Heat.Type + Living.Area + Bedrooms +                 Bathrooms + Rooms,
        house,
        method = "lm",
        trControl = tr_ctrl
        )

result_back = cv_back$results


# forward
cv_fwd = train(
        Price ~ Living.Area + Land.Value + Bathrooms + 
                Waterfront + New.Construct + Heat.Type + Lot.Size + Central.Air + 
                Age + Rooms + Bedrooms,
        house,
        method = "lm",
        trControl = tr_ctrl
        )

result_fwd = cv_fwd$results

# stable
cv_stable = train(
        Price ~ Waterfront + Land.Value + Living.Area + Bathrooms, house,
        method = "lm",
        trControl = tr_ctrl
        )

result_stable = cv_stable$results

results <-resamples(list(backward = cv_back, forward = cv_fwd, stable = cv_stable))
```


In [Analysis](@sec-analysis) and [Model Stability](@sec-stability) sections, three valid models are found, namely, backward stepwise regression model (backward model), forward stepwise regression model (forward model), and stable model. 

In this section, we further compare their in-sample performances and out-of-sample performances.

## In-sample performaces

In-sample perfomances are evaluated by calculating $r^2$, and adjusted $r^2$ values of the three models within the dataset (Table \ref{tab:rsquare}). We found that backward and forward models have the same $r^2$ (`r round(rsquare_back, 3)`) and adjusted $r^2$ values (`r round(rsquare_adj_back, 3)`). The stable model have slightly lower $r^2$ and adjusted $r^2$ values than the two stepwise models, but very similar (`r round(rsquare_stable, 3)` and `r round(rsquare_adj_stable, 3)` respectively).

\begin{table}[!htbp]
  \begin{center}
    \begin{tabular}{ |c|c|c| } 
      \hline
      Models & $r^2$ & Adjusted $r^2$ \\ 
      \hline
      Backward & `r round(rsquare_back, 3)` & `r round(rsquare_adj_back, 3)` \\ 
      \hline
      Forward & `r round(rsquare_fwd, 3)` & `r round(rsquare_adj_fwd, 3)` \\ 
      \hline
      Stable & `r round(rsquare_stable, 3)` & `r round(rsquare_adj_stable, 3)` \\ 
      \hline
    \end{tabular}
    \caption{$r^2$ and adjusted $r^2$}
    \label{tab:rsquare}
  \end{center}
\end{table}

## Out-of-sample performances

A $10$-fold cross validation shows that backward and forward models have slightly lower average MAE (Mean Absolute Error) and RMSE (Root Mean Square Error) than the stable model (Table \ref{tab:error}).

\begin{table}[!htbp]
  \begin{center}
    \begin{tabular}{ |c|c|c| } 
      \hline
      Models & RMSE & MAE \\ 
      \hline
      Backward & `r format(result_back[1,2], nsmall = 3)` & `r format(result_back[1,4], nsmall = 3)` \\ 
      \hline
      Forward & `r format(result_fwd[1,2], nsmall = 3)` & `r format(result_fwd[1,4], nsmall = 3)` \\ 
      \hline
      Stable & `r format(result_stable[1,2], nsmall = 3)` & `r format(result_stable[1,4], nsmall = 3)` \\ 
      \hline
    \end{tabular}
    \caption{RMSE and MAE}
    \label{tab:error}
  \end{center}
\end{table}

Figures (Figure \ref{fig:rmse}, Figure \ref{fig:mae}) of RMSE and MAE further confirm it, though majority parts of the range of RMSE and MAE among the three models overlap. 


```{r fig.cap="\\label{fig:rmse}RMSE", out.width = "150px"}
ggplot(results, metrics = "RMSE") +
  labs(y = "")
```

```{r fig.cap="\\label{fig:mae}MAE", out.width = "150px"} 
ggplot(results, metrics = "MAE") +
  labs(y = "")
```


# Discussion, Conclusion, and Limitation

## Discussion

In-sample performance evaluation indicates that the backward and forward models explain the data slightly more than the stable model due to their higher $r^2$ and adjusted $r^2$ values.

Out-of-sample performance evaluation further shows that the backward and forward models may be slightly more accurate than the stable model due to their lower RMSE and MAE values.

However, as mentioned in the [Model Stability](@sec-stability) section, the backward and forward models with $14$ parameters are not stable. In comparison, the stable model is selected in $78\%$ of bootstrap resamples.

Meanwhile, though the stable model has higher $r^2$ and adjusted $r^2$ values than the backward and forward models, it still manages to explain approximately $`r round(rsquare_stable*100, 2)` \%$ (according to its $r^2$ value) or $`r round(rsquare_adj_stable*100, 2)` \%$ (according to its adjusted $r^2$ value) of the total variation in the house price data. This is reasonable. Furthermore, the stable model does not have much higher error rates than the backward and forward models. 

## Conclution

Balancing stability, in-sample performances and out-of-sample performances, we conclude that the stable model may be the best among the four models.

## Limitation

As mentioned, when compared to the backward and forward models, the stable model is more stable. However, it has relatively slightly lower $r^2$ and adjusted $r^2$ values, as well as relatively higher error rates (RMSE and MAE). To some extend, it is a compromise between stability and accuracy.

If more information about the dataset is provided, a domain knowledge expert may make better judgement on which model to choose.





