---
title: "House Prices in New York"
author: "Group CC901E2"
institute: "The University of Sydney"
output: html_notebook
format:
  revealjs:
    self-contained: true
    fig-format: retina
    toc: true
    toc-depth: 1
    number-sections: false
    code-line-numbers: false
    slide-number: c
    scrollable: true
    pdf-max-pages-per-slide: 1
    history: false # don't add each slide to browser history
    
execute:
  echo: false
  warning: false # do not show warning in the chunk
  message: false # do not show message in the chunk
---

```{r message=FALSE}
# Load packages here
library(tidyverse)
library(ggplot2)
library(ggfortify)
library(dplyr)
library(ISLR)
library(sjPlot)
library(leaps)
library(GGally)
library(caret)
library(mplot)
library(lmSubsets)
```

```{r, message=FALSE}
# The original data is loaded here
#| output: FALSE
house = read.delim("housing-prices-ge19.txt", 
                    header = TRUE)
# convert data types
house = house %>% mutate (
  Waterfront = as.factor(Waterfront),
  New.Construct = as.factor(New.Construct),
  Central.Air = as.factor(Central.Air),
  Fuel.Type = as.factor(Fuel.Type),
  Heat.Type = as.factor(Heat.Type),
  Sewer.Type = as.factor(Sewer.Type),
  Test = as.factor(Test)
)

# Remove the Test column, because we do not have the metadata
house = house %>%
  select(-Test)

# # Trying to filter some of the variables 
# house_clean = house %>% 
#   dplyr::mutate(
#     Lot.Size = case_when(
#     Lot.Size < 1000 ~ NA_real_,
#      is.na(Lot.Size) ~ NA_real_,
#      TRUE ~ Lot.Size
#     )
#   ) %>% drop_na(Lot.Size)
# summary(house_clean)

# clean names
# house = janitor::clean_names(house)
```



# Stable Model

```{r, echo=FALSE}
load("house_main.RData")
```


```{r, fig.fullwidth=TRUE}
# plot(vis.h, which = "vip", interactive = FALSE, legend.position = "bottom")
```
Through the variable inclusion plot, it can be found that  `Land.Value`, `Living.Area`, `Bathrooms`, `New.Construct`, `Waterfront` are the five most important variables for predicting `Price`, while the `Sewer.TypePublic` and all levels of `FuelType` do not provide any useful information, since they are included in models by chance.


## Model stability Plot {.smaller}

-   There appears to be dominant models in models of size **two, three, and five** (including the intercept), as demonstrated by one of the circles being substantially larger than the other circles with models of the same size. 
-   Also, the stepwise model of size 14 is not stable because of their small circle size. 
-   Finally, it can be found that the simple reversion models is stable, as it is always selected in bootstrap resamples.

```{r, fig.fullwidth=TRUE}
plot(vis.h, highlight = "Living.Area", which = "boot", interaction = F, max.circle = 10, seed = 2022, legend.position = "bottom")
```


```{r}
# print(vis.h, min.prob = 0.5)
```

## Conclusion: 
The $Price \sim Waterfront1 + Land.Value + Living.Area + Bathrooms$ model is selected in 78% of bootstrap resamples, which strikes the balance between accuracy and stability.

